\chapter*{Tutorial 1}
\section{Truth tables and semantics}

\begin{exercise}
Convert any formula into an equivalent formula which uses only symbols $\wedge$ and $\neg$.
\end{exercise}

\begin{exercise}
Write a formula which satisfies the following condition
\begin{align*}
p_1p_2\dots p_n  \\
+ \\
q_1q_2 \dots q_n &  \\
= \\
r_1r_2 \dots r_n
\end{align*}
View $p_1,\dots,p_n,q_1,\dots,q_n,r_1,\dots,r_n$ as propositions.
\end{exercise}

\begin{exercise}
Use propositional logic to answer the following question:

You are trapped in a room. There are two doors. Either the doors lead to an exit or to a lion (note that both leading to an exit or to a lion are also possible). In Door 1, it is written ``This door is exit and other door leads to lion". In Door 2, it is written ``One of the rooms lead to exit, the other to a lion". You are told that only one of the written statements are true and the other false. Which door would you choose?
\end{exercise}

\begin{exercise}[Pigeon hole principle]
Consider $n+1$ pigeons and $n$ holes. Let proposition $p_{i,j}$ denote the fact that the $i^{th}$ pigeon is in the $j^{th}$ hole.  Write a propositional logic formula to encode the pigeon hole principle. 
\end{exercise}

\begin{exercise}
Show that the following statments are equivalent.
\begin{enumerate}
\item $\alpha_1,\alpha_2,\dots,\alpha_n \models \beta$
\item \true $~ \models \Big(\alpha_1 \ifthen \big(\alpha_2 \ifthen \dots (\alpha_n \ifthen \beta)\big)\Big)$
\item \true $~\models (\alpha_1 \wedge \alpha_2 \wedge \dots \alpha_n) \ifthen \beta$
\end{enumerate}
\end{exercise}

\section{Natural deduction}
\begin{exercise}
Derive natural deduction proofs for the following
\begin{enumerate}
\item \true $~\proves \alpha \vee \neg \alpha$.
\item $\neg (\alpha \wedge \beta) \dashv \proves \neg \alpha \vee \neg \beta$.
\item \true  $~\proves \alpha \ifthen (\beta \ifthen \alpha)$
\item $\alpha \vee \beta, \alpha \vee \neg \beta \proves \alpha$
\item $\neg \alpha \proves \alpha \ifthen \beta$
\item $\neg \alpha \proves \neg (\alpha \wedge \beta)$
\end{enumerate}
\end{exercise}


\begin{exercise}
Show that the following are equivalent
\begin{enumerate}
\item $\alpha_1, \alpha_2, \dots, \alpha_n \proves \beta$
\item \true $~\proves \Big(\alpha_1 \ifthen \big(\alpha_2 \ifthen (\dots (\alpha_n \ifthen \beta))\big)\Big)$
\item \true $~\proves (\alpha_1 \wedge \alpha_2 \wedge \dots \alpha_n) \ifthen \beta$
\end{enumerate}
\end{exercise}

\begin{exercise}
Answer the following questions
\begin{enumerate}
\item Let $\Psi$ be a formula over only the proposition $p$. Assume that $p \proves \psi$ and $\neg p \proves \psi$. Show that,  \true $~\proves \psi$.
\item This is an extension of the previous one. Let $\psi$ be a formula over proposition $P$. Let $\Gamma$ be the set of all formulas over $P$. If $\alpha \proves \psi$ for all $\alpha \in \Gamma$, then $\true ~\proves \psi$.
\end{enumerate}
\end{exercise}
