\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{enumitem}
\setlist{nosep}

\begin{document}

\begin{center}
    {\LARGE \textbf{Curriculum Vitae}} \\
    {\large \textbf{A. V. Sreejith}} \\
    \href{mailto:sreejith@iitgoa.ac.in}{sreejith@iitgoa.ac.in} | \href{https://yourwebsite.com}{yourwebsite.com} \\
    Associate Professor, School of Mathematics and Computer Science, IIT Goa
\end{center}

\vspace{0.5cm}

\section*{Professional Summary}
Experienced researcher and educator specializing in theoretical computer science, formal methods, and data science. Award-winning teaching professional with a proven track record in research funding, curriculum development, and mentoring students.

\section*{Research Interests}
\begin{itemize}
    \item \textbf{Theoretical Computer Science:} Automata theory, Descriptive complexity, Logic in computer science.
    \item \textbf{Formal Methods and Verification:} Model checking, Probabilistic systems, Formal methods in machine learning.
    \item \textbf{Data Science:} High-dimensional data analysis, Formal approaches to learning algorithms.
\end{itemize}

\section*{Current Position}
\textbf{Associate Professor}, School of Mathematics and Computer Science \\
Indian Institute of Technology (IIT) Goa \hfill August 2021 -- Present \\
\emph{Responsibilities:} Teaching undergraduate and graduate courses, curriculum development, research supervision, and administrative leadership.

\section*{Education}
\begin{itemize}
    \item \textbf{Ph.D. in Mathematics}, Institute of Mathematical Sciences (IMSc), Chennai \\
    Advisor: Prof. Kamal Lodaya | Thesis: ``Regular Quantifiers in Logics'' \hfill August 2007 -- September 2013
    \item \textbf{M.Tech in Computer Science}, Indian Institute of Technology (IIT) Madras \hfill August 2005 -- May 2007
    \item \textbf{B.Tech in Computer Science}, College of Engineering, Thiruvananthapuram \hfill August 2000 -- June 2004
\end{itemize}

\section*{Research Positions Held}
\begin{itemize}
    \item \textbf{Professeur Invit\'e}, University of Bordeaux, France \hfill April -- July 2024
    \item \textbf{Postdoctoral Researcher}, University of Warsaw \hfill 2016 -- 2017 \\
    Mentor: Prof. Mikołaj Bojańczyk
    \item \textbf{Postdoctoral Researcher}, Laboratoire d'Informatique Algorithmique: Fondements et Applications (LIAFA), Paris, France \hfill 2014 -- 2015 \\
    Mentor: Prof. Thomas Colcombet
    \item \textbf{Postdoctoral Researcher}, Tata Institute of Fundamental Research (TIFR), Mumbai \hfill 2013 -- 2014
    \item Researcher, University of Tübingen, Germany \hfill 2012 \\
    Mentor: Prof. Klaus-Jörn Lange
\end{itemize}

\section*{Awards and Honors}
\begin{itemize}
    \item \textbf{Professeur Invit\'e}, University of Bordeaux (2023-2024).
    \item \textbf{ACM India Doctoral Dissertation Award (Honorable Mention)}, 2014.
\end{itemize}

\section*{Research Funding}
\begin{itemize}
    \item \textbf{CEFIPRA Project:} Formal Verification of Adaptive Control Algorithms \\
    Collaborators: Dr. Vincent Penelle (University of Bordeaux), Dr. Meenakshi D'Souza (IIIT Bangalore).
    \item \textbf{MATRICS Project:} Probabilistic Pushdown Automata, DST India.
    \item \textbf{MoES Project:} Remote Sensing for River Discharge Estimation.
\end{itemize}

\section*{Teaching Experience}
\textbf{Indian Institute of Technology Goa (IIT Goa)} \\
\begin{itemize}
    \item \textbf{Undergraduate Courses:} Data Structures and Algorithms, Compilers, Computer Networks.
    \item \textbf{Graduate Courses:} Advanced Algorithms, Formal Methods in Machine Learning, Foundations of Theoretical Computer Science.
    \item \textbf{International Teaching:} Formal Methods in Machine Learning (LaBRI, University of Bordeaux, 2024).
\end{itemize}

\section*{Publications}
\textbf{Selected Peer-Reviewed Journal Articles}
\begin{enumerate}
    \item ``Algebraic Characterizations and Block Product Decompositions...'', \emph{Journal of Computer and System Sciences (JCSS)}, 2023.
    \item ``Undecidability of a Weak Version of MSO+U'', \emph{Logical Methods in Computer Science (LMCS)}, 2020.
\end{enumerate}

\textbf{Conference Proceedings}
\begin{enumerate}
    \item ``One Deterministic Counter Automata'', \emph{FSTTCS}, 2023.
    \item ``First-Order Logic and Infinitary Quantifiers...'', \emph{LICS}, 2019.
\end{enumerate}

\section*{Supervision}
\textbf{Ph.D. Students:} Prince Mathew (January 2019 -- Present) \\
\textbf{M.Tech Thesis:} Rashika N (2024), Sumitra (2024), Ankita Chand (2023).

\section*{Administrative Roles}
\begin{itemize}
    \item Program Chair, Computer Science and Engineering, IIT Goa.
    \item Chair, SSAC; Head of IT Infrastructure, IIT Goa.
    \item Program Chair, ICLA 2023.
\end{itemize}

\end{document}
