 % LaTeX resume using res.cls
\documentclass[margin]{res}
\usepackage{amsfonts}

%\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
%\usepackage{newcent}   % uses new century schoolbook postscript font 
\setlength{\textwidth}{5.1in} % set width of text portion
\usepackage{etaremune}
\usepackage{hyperref}
\hypersetup{
hidelinks
}

\begin{document}

% Center the name over the entire width of resume:
 \moveleft.5\hoffset\centerline{\Large \bf A. V. Sreejith} 
% Draw a horizontal line the whole width of resume:
 \moveleft\hoffset\vbox{\hrule width\resumewidth height 1pt}\smallskip
% address begins here
% Again, the address lines must be centered over entire width of resume:
\moveleft.5\hoffset\centerline{Indian Institute of Technology Goa}
\moveleft.5\hoffset\centerline{GEC campus, Farmagudi, Goa - 403401}
\moveleft.5\hoffset\centerline{E-mail: sreejithav@iitgoa.ac.in}
\moveleft.5\hoffset\centerline{ \href{http://www.iitgoa.ac.in/~sreejithav/} {\tt{www.iitgoa.ac.in/$\sim$sreejithav/}}}
%\moveleft.5\hoffset\centerline{Phone: +918939601670}

 
\begin{resume}
\section{Publications (Aug 2021 - July 2022)} 
\begin{enumerate}
\item Bharat Adsul, Saptarshi Sarkar, A. V. Sreejith, ``\emph{Block Product Decompositions for First Order Logic and its Infinitary Quantifier Extensions}'', Journal of Computer and System Sciences (under process)
\item Bharat Adsul, Saptarshi Sarkar, A. V. Sreejith, ``\emph{First-Order logic and its Infinitary Quantifier Extensions over Countable Words}", International Symposium on Fundamentals of Computation Theory (FCT), pages 39-52, 2021.
\end{enumerate}
\section{Publications before the period under consideration}
\begin{enumerate}
\item Gaurav, K., Métivier, F., A V Sreejith, Sinha, R., Kumar, A., and Tandon, S. K.: ``\emph{Coupling threshold theory and satellite-derived channel width to estimate the formative discharge of Himalayan foreland rivers}'', Earth Surf. Dynam., 9, 47-70, https://doi.org/10.5194/esurf-9-47-2021, 2021
\item M. Bojanczyk, L. Daviaud, B. Guillon, V. Penelle, A. V. Sreejith ``\emph{Undecidability of a weak version of MSO+U}", Logical Methods in Computer Science (LMCS), volume 16:1, 2020.
\item Bharat Adsul, Saptarshi Sarkar, A. V. Sreejith, ``\emph{Block products for algebras over countable words and applications to logic}", Logic in Computer Science (LICS), pages 1-13, 2019.
\item Kamal Lodaya, A. V. Sreejith, ``\emph{Two variable first order logic with counting quantifiers: complexity results}'', Developments in Language Theory (DLT), LNCS 10396, pages 260 - 271, 2017.
\item Amaldev Manuel, A. V. Sreejith, ``\emph{Two variable logic over countable linear orderings}", Mathematical Foundations of Computer Science (MFCS), LIPIcs 58, pages 66:1-66:13, 2016. 
\item Thomas Colcombet, A. V. Sreejith, ``\emph{Limited Set quantifiers over Countable Linear Orderings}", Proceedings of the Automata, Languages, and Programming - 42nd International Colloquium (ICALP), volume 9135 of LNCS, pages 146-158, 2015. 
\item Kamal Lodaya, A. V. Sreejith, ``\emph{Counting quantifiers and linear arithmetic on word models}" with Kamal Lodaya in 14th Asian Logic Conference (ALC), 2014.
\item V. Arvind, S. Raja, A. V. Sreejith, ``\emph{On lower bounds for multiplicative circuits and linear circuits in noncommutative domains}", 9th International Computer Science Symposium in Russia (CSR), volume 8476 of LNCS, pages 65-76, 2014.
\item Andreas Krebs, A. V. Sreejith, ``\emph{Non-definability of Languages by Generalized First-order Formulas over (N, +)}", Proceedings of the 27th Annual IEEE Symposium on Logic in Computer Science (LICS),  pages 451-460, 2012.
\item A. V. Sreejith, ``\emph{Expressive Completeness for LTL With Modulo Counting and Group Quantifiers}", Electronic Notes in Theoretical Computer Science, ENTCS 278, pages 201-214, 2011.
\item Kamal Lodaya, A. V. Sreejith, ``\emph{LTL can be more succinct}", Proceedings of Automated Technology for Verification and Analysis (ATVA), LNCS 6252, 21-24, 2010.
\end{enumerate}

\end{resume}
\end{document}




